---
title: "Logistic Regression"
author: "Howard Baek"
date: "7/14/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Plot 1)

Daily at different 4 latitudes. X-axis is upwelling index (continuous); Y-axis is 0, 1 (whether categorized as upwelling via your algorithm)

The closest: SST erddap data with latitude 42.125. there's no 42.000 latitude

## Libraries
```{r}
library(tidyverse)
library(lubridate)
library(scales)
library(factoextra)
library(raster)
library(tmap)
library(png)
library(gridExtra)
library(dendextend)
library(DescTools)
library(imageryML)
library(plotly)
library(ggmap)
library(rerddap)

theme_set(theme_light())
```

## Load data

```{r}
df_raw <- read_csv("data/erddap_2000-01-01_2021-06-01_lat_40.625_50.625_lon_229.875_236.625.csv")

# Process data
df_processed <- df_raw %>% 
  # Get rid of miscellaneous zlev in first row
  slice(-1) %>% 
  # zlev is a column of zeroes, so get rid of that
  dplyr::select(-zlev) %>% 
  # Convert into date
  mutate(time = ymd_hms(time)) %>% 
  # Set column names
  rename(date = time,
         lat = latitude,
         lon = longitude) %>% 
  # Convert date column to Date type
  mutate(date = as.Date(date),
         lat = as.numeric(lat),
         lon = as.numeric(lon),
         sst = as.numeric(sst))

# mask out Puget Sound, Strait of Juan de Fuca and Georgia Strait
masks <- list(c(235.4488, 236.884, 47.87651, 50.13138),
              c(232.2913, 233.8987, 50.28689, 51.60871),
              c(234.4154, 235.9654, 49.04283, 50.09251))

for (m1 in masks) {
  # index of Puget Sound, Strait of Juan de Fuca or Georgia Strait
  mask_loc <- df_processed$lat <= m1[4] & df_processed$lat >= m1[3] &
    df_processed$lon <= m1[2] & df_processed$lon >= m1[1]
  # Change to NA
  df_processed$sst[mask_loc] <- NA
}

# Create df for specific latitude: 42.125N and -125.125 longitude
df_processed_42_125 <- df_processed %>% 
  filter(lat == 42.125) %>% 
  mutate(lon = lon - 360) 

# Automatic Detection Method
# Find SST values next to land (sst_coast_X) and check if upwelling
is_upwelling <- df_processed_42_125 %>% 
  group_by(date, lat) %>% 
  # 1 longitude tick away from land
  mutate(sst_coast_1 = last(na.omit(sst)),
         # 8 longitude ticks away from land (2 degrees in longitude)
         sst_coast_2 = nth(na.omit(sst), -9),
         # 12 longitude ticks away from land (3 degrees in longitude)
         sst_coast_3 = nth(na.omit(sst), -13)) %>% 
  # Find difference between pixels
  # Threshold: 0.15
  summarize(is_upwelling_1_2 = sst_coast_2 - sst_coast_1 > 2,
            is_upwelling_1_3 = sst_coast_3 - sst_coast_1 > 2) %>% 
  ungroup() %>% 
  group_by(date, lat) %>% 
  # Check if any upwelling in each latitude
  # first() because I want to return one row for each lat
  summarise(is_upwelling_total = first(is_upwelling_1_2 | is_upwelling_1_3)) %>% 
  ungroup()

is_upwelling
```

is_upwelling contains daily date, latitude, and is_upwelling (TRUE, FALSE)


## Get Upwelling Index data

Upwelling Index, 42N 125W, 6-hourly: erdUI426hr

```{r}
ui <- tabledap("erdUI426hr") %>% 
  as_tibble() %>% 
  dplyr::select(-station_id) %>%
  # Convert date column to Date type
  transmute(date = as.Date(time),
            upwelling_index,
            lat = as.numeric(latitude),
            lon = as.numeric(longitude)) %>% 
  filter(date >= "2000-01-01") %>% 
  group_by(date, lon, lat) %>% 
  # Average 6 hour upwelling index to get daily index
  summarise(upwelling_index_daily_avg = mean(upwelling_index), .groups = "drop")

ui
```



## Merge is_upwelling with ui data

`is_upwelling` only has data until 2021-06-01
`ui` has data until 2021-06-30

so ui has 30ish more rows

```{r}
final_df <- ui %>% 
  inner_join(is_upwelling, by = "date") %>% 
  select(date, lat = lat.x, lon, upwelling_index_daily_avg, is_upwelling_total) %>% 
  mutate(is_upwelling_total = as.numeric(is_upwelling_total))

final_df
```


## Logistic Regression

```{r}
logistic_fit <- glm(is_upwelling_total ~ upwelling_index_daily_avg, data = final_df, family = "binomial")

summary(logistic_fit)
```

Shows that daily avg of upwelling index has a statistically significant, positive relationship with whether upwelling or not.

