---
title: "Untitled"
author: "Howard Baek"
date: "7/4/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rerddap)
```


## Upwelling Index

```{r}
ui_search <- ed_search(query = "upwelling index", which = "tabledap",
                       url = "https://coastwatch.pfeg.noaa.gov/erddap/")
ui_search

ui_info <- info("erdUI486hr")
ui_info

# Upwelling Index, 48N 125W, 6-hourly
fourtyeight <- tabledap("erdUI486hr")
fourtyeight
```


The result of `ed_search` gives us a df with title column, which is the Dataset title on ERDDAP 

I checked and found that `fourtyeight` gives the same data table as that from https://coastwatch.pfeg.noaa.gov/erddap/tabledap/erdUI486hr.html

This confirms the validity of the `tabledap()` function!



## SST, Daily Optimum Interpolation

```{r}
sst_search <- ed_search(query = "sea surface temperature Optimum Interpolation", which = "griddap", url = "https://coastwatch.pfeg.noaa.gov/erddap/")

sst_search <- sst_search$info

sst_search %>% 
  filter(str_detect(title, "SST|Daily Optimum Interpolation"))
```


SST, Daily Optimum Interpolation (OI) dataset_id: 

```{r}
sst_info <- info("ncdcOisst21Agg")

sst_df <- griddap("ncdcOisst21Agg", latitude = c(40.625, 45.625), longitude = c(229.875, 236.625), time = c("1982-01-01", "2021-06-01"), fields = "sst")$data

beepr::beep("mario")
```

I checked and found that `sst_df` gives the same data table as that from https://coastwatch.pfeg.noaa.gov/erddap/griddap/ncdcOisst21Agg.htmlTable?sst%5B(2021-06-19T12:00:00Z)%5D%5B(0.0)%5D%5B(25.375):(58.375)%5D%5B(218.125):(251.125)%5D&.draw=surface&.vars=longitude%7Clatitude%7Csst&.colorBar=%7C%7C%7C16%7C32%7C&.bgColor=0xffccccff

This confirms the validity of the `griddap()` function!
